Gyrochronal and Lithium Ages for Stars and Planets Observed by Kepler

### 2023-06-05 paper/ms.tex
update ms

Theoretical predictions for planet evolution in the Kepler sample are
most important in the context of mini-Neptunes and super-Earths, since
these are the most abundant planets.  Pertinent predictions include:
{\it a)} that mini-Neptunes should cool and contract, roughly
following a $R_p \propto t^{-0.1}$ scaling \citep{Gupta_2019}; {\it
b)} that large and close-in mini-Neptunes at the edge of the
photoevaporation desert should shrink over the atmospheric loss
timescale, $t_{\rm loss}$ \citep{Owen_Lai_2018}; {\it c)} that the
abundance of 4-8\,R$_\oplus$ planets with masses of 5-30\,M$_\oplus$
should decrease over $t_{\rm loss}$---the upper-envelope of the
mini-Neptune size distribution should corresponding decrease,
e.g.~from $\approx$4.5\,R$_\oplus$ at $\approx$200\,Myr to
$\approx$3.5\,R$_\oplus$ at $\approx$700\,Myr
\citep[e.g.][]{Rogers_2021}; {\it d)} that the super-Earth to
mini-Neptune ratio, $\theta$, should increase over $t_{\rm loss}$
\citep[e.g.][]{Rogers_2021} {\bf BY HOW MUCH?}); {\it e)} that the
upper boundary of the super-Earth population should increase over the
same timescale as $\theta$ is increasing, since more massive
super-Earth cores should be able to retain their envelopes for longer
than less massive cores.

@@

As our default source for stellar temperatures and surface gravities, we adopt
the values from the Gaia-Kepler Stellar Properties Catalog (GKSP;
\citealt{Berger_2020a_catalog}).
The GKSP parameters are available for stars with ``AAA'' 2MASS photometry,
measured parallaxes in Gaia DR2,  and g-band photometry from either the KIC or
the Kepler-INT survey.
The parameters themselves were derived using \texttt{isoclassify} (CITE
Huber2017) to interpolate over the MIST isochrone grids (CITE Choi2016,
Dotter2016), given the SDSS $g$ and 2MASS $K_{\rm s}$ photometry, the Gaia DR2
parallexes, and $V$-band extinction from the (CITE) Green2019 reddening map.
The resulting stellar parameters are available for
\fracstarswithprotwithbtwenty\ of the \nstarswithprot\ Kepler stars
with rotation periods.
For the remaining \fracstarswithprotwithoutbtwenty\ of stars that lack
temperatures and surface gravities, we adopt the values 
from the Gaia DR3 General Stellar Parametrizer from Photometry (GSP-Phot; CITE).
GSP-Phot uses the low-resolution Gaia BP and RP spectra, mean $G$ magnitudes,
and parallaxes, in concert with the PARSEC isochrones (CITE Tang2014, Chen2015,
Pastorelli2020) to infer the stellar effective temperatures, surface gravities,
metallicities, and line-of-sight extinctions.
In the planet sample,
\frackoisnofpwithprotwithbtwenty\ of the
the non-false-positive KOIs with rotation periods have parameters from
\citet{Berger_2020a_catalog}.

@@

>>>>

Theoretical predictions for planet evolution in the Kepler sample are
most important in the context of mini-Neptunes and super-Earths, since
these are the most abundant planets.  Pertinent predictions include:
{\it a)} that mini-Neptunes should cool and contract, roughly
following a $R_p \propto t^{-0.1}$ scaling \citep{Gupta_2019}; {\it
b)} that large and close-in mini-Neptunes at the edge of the
photoevaporation desert should shrink over the atmospheric loss
timescale, $t_{\rm loss}$ \citep{Owen_Lai_2018}; {\it c)} that the
abundance of 4-8\,R$_\oplus$ planets with masses of 5-30\,M$_\oplus$
should decrease over $t_{\rm loss}$---the upper-envelope of the
mini-Neptune size distribution should corresponding decrease,
e.g.~from $\approx$4.5\,R$_\oplus$ at $\approx$200\,Myr to
$\approx$3.5\,R$_\oplus$ at $\approx$700\,Myr
\citep[e.g.][]{Rogers_2021}; {\it d)} that the super-Earth to
mini-Neptune ratio, $\theta$, should increase over $t_{\rm loss}$
\citep[e.g.][]{Rogers_2021}, with the mini-Neptune abundance
decreasing by at least a factor of two or three, and the super-Earth
abundance correspondingly increasing {\bf BY HOW MUCH?}); {\it e)} that the
upper boundary of the super-Earth population should increase over the
same timescale as $\theta$ is increasing, since more massive
super-Earth cores should be able to retain their envelopes for longer
than less massive cores.

@@

As our default source for stellar temperatures and surface gravities, we adopt
the values from the Gaia-Kepler Stellar Properties Catalog (GKSP;
\citealt{Berger_2020a_catalog}).
The GKSP parameters are available for stars with ``AAA'' 2MASS photometry,
measured parallaxes in Gaia DR2,  and g-band photometry from either the KIC or
the Kepler-INT survey.
The parameters themselves were derived using \texttt{isoclassify} (CITE
Huber2017) to interpolate over the MIST isochrone grids (CITE Choi2016,
Dotter2016), given the SDSS $g$ and 2MASS $K_{\rm s}$ photometry, the Gaia DR2
parallexes, and $V$-band extinction from the (CITE) Green2019 reddening map.
The resulting stellar parameters are available for
\fracstarswithprotwithbtwenty\ of the \nstarswithprot\ Kepler stars
with rotation periods.

%For the remaining \fracstarswithprotwithoutbtwenty\ of stars that lack
%temperatures and surface gravities, we adopt the values 
%from the Gaia DR3 General Stellar Parametrizer from Photometry (GSP-Phot; CITE).
%GSP-Phot uses the low-resolution Gaia BP and RP spectra, mean $G$ magnitudes,
%and parallaxes, in concert with the PARSEC isochrones (CITE Tang2014, Chen2015,
%Pastorelli2020) to infer the stellar effective temperatures, surface gravities,
%metallicities, and line-of-sight extinctions.
%In the planet sample,
%\frackoisnofpwithprotwithbtwenty\ of the
%the non-false-positive KOIs with rotation periods have parameters from
%\citet{Berger_2020a_catalog}.

For the remaining \fracstarswithprotwithoutbtwenty\ of stars that lack
temperatures and surface gravities, we adopt the values reported by
\citet{Santos2019} and \citet{Santos2021}, which for these cases
primarily derive from the \citet{Mathur_2017} DR25 Kepler Stellar
Properties Catalog, and are mostly derived from photometry.  In the
planet sample, \frackoisnofpwithprotwithbtwenty\ of the the
non-false-positive KOIs with rotation periods have parameters from
\citet{Berger_2020a_catalog}, and the remainder similarly are drawn
from DR25.

@@

\subsection{Evolutionary trends in time}

{\it Kelvin-Helmholtz cooling}
$R_p \propto t^{-0.1}$ scaling \citep{Gupta_2019}

{\it Early carving of the photoevaporation desert}
\citep{Owen_Lai_2018}

{\it Rapidly decreasing abundance of big puffy planets}
Kepler not enough?

{\it Mini-Neptunes turning into Super-Earths}
\citep[e.g.][]{Rogers_2021}

{\it Boundary of the Super-Earth population moving up in time}
\citep{David_2021}
